---
title: "Live coding exercise - Pivoting"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
execute:
  echo: true
  warning: false
bibliography: references.bib
---

```{r}
library(tidyverse)
library(gapminder)
```

# tidyr: pivoting

## Example: Waste characterisation data

The data for this exercise was generated as part of a course taught at ETH Zurich, Switzerland (https://rbtl-fs22.github.io/website/). A repository that contains all generated data is in preparation.

The dataset contains the results of a waste characterisation study of public bins in Zurich, Switzerland. Waste was separated and weighed in kg for the following categories:

-   paper
-   PET
-   metals
-   glass
-   others

Additional metadata on location and type of bin is available.

```{r}



```

```{r}


```

## Example: Gapminder

The following exercise is taken from [@zimmerman2019].

```{r}

gapminder_wide <- read_csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/main/data/gapminder_wide.csv")

```

```{r}


```

### From wide to long

```{r}


```

**Your turn**

1.  Using `gapminder_long`, calculate the mean life expectancy, population, and gdpPercap for each continent. **Hint:** use the `group_by()` and `summarize()` functions we learned in the `dplyr` lesson.
2.  Assign the created dataframe to an object with the name `gapminder_long_summary`

```{r}

```

### From long to intermediate format

```{r}


```
