---
title: "Untitled"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Project report

Thank you for uploading your identified data to the course material. I have created this file for you to work on your data using the code you have learned as part of the course. I have added some data cleaning steps as the first code-chunk. I hope this will help you work with your data more efficiently.

It is up to you which analysis you would like to perform with your data. I am available to support you and answer any questions you may have.

```{r}
# here I load all R packages that are part of the tidyverse collection of R packages
# you can use this instead of loading each package individually

library(tidyverse)

```

# Data import

```{r}
emissions <- read_csv("data/energy_co2 _emissions_emerging_economies_2010s.csv")

```

# Data cleaning

Your data is already in a good format. I have only changed the capitalization of the variable names, but that is out of preference and not necessary as they are following the same pattern.

```{r}
emissions_tidy <- emissions |>
  rename_all(tolower) 
```

# Data transformation

I have filtered the data for 2019 and prepare a summary by country and fueltype, but there are many other options you have here for working with the data.

```{r}

emissions_tidy_sum <- emissions_tidy |> 
  filter(year == 2019) |> 
  group_by(country, fueltype) |> 
  summarise(
    sum_co2_mt = sum(co2_mt)
  )

```

# Data visualisation

```{r}

emissions_tidy_sum |> 
  filter(sum_co2_mt < 100) |> 
  # it is possible to use the pipe to use the data directly in 
  # the ggplot() function for plotting
  ggplot(mapping = aes(x = fueltype, 
                       y = sum_co2_mt)) +
  geom_boxplot() 

```

```{r}

emissions_tidy_sum |> 
  filter(sum_co2_mt > 100) |> 
  ggplot(mapping = aes(x = fueltype,
                       y = sum_co2_mt)) +
  geom_boxplot() 

```

## 
