---
title: "World Bank - World Development Indicators"
subtitle: "Youth Unemployment"
format: revealjs
editor: source
editor_options: 
  chunk_output_type: console
self-contained: true
execute:
  eval: true
  echo: true
  warning: false
---

```{r}
#| echo: false

library(tidyverse)
library(WDI)
library(countrycode)

```

# Introduction

## World Bank - World Development Indicators (WDI)

> "World Development Indicators (WDI) is the World Bank's premier compilation of cross-country comparable data on development." - [World Development Indicators Website](https://datatopics.worldbank.org/world-development-indicators/)

> "DataBank is an analysis and visualisation tool that contains collections of time series data on a variety of topics." - [World Bank DataBank Tool](https://databank.worldbank.org/home.aspx)

## WDI R Package

> "The WDI package allows users to search and download data from over 40 datasets hosted by the World Bank, including the World Development Indicators ('WDI'), International Debt Statistics, Doing Business, Human Capital Index, and Sub-national Poverty indicators." - [WDI R Package Documentation](https://vincentarelbundock.github.io/WDI/index.html) by ___

## Selected indicator - "Youth Unemployment"

> Share of youth not in education, employment or training (NEET) is the proportion of young people who are **not in education, employment, or training** to the population of the corresponding age group: **youth (ages 15 to 24); persons ages 15 to 29; or both age groups**.

# Methods

## Data search

```{r}

WDIsearch("electricity") |> 
  as_tibble()

WDIsearch("gender") |> 
  as_tibble()

WDIsearch("share of youth") |> 
  as_tibble()
```

## Define indicators as vector

```{r}

selected_indicators <- c("SL.UEM.NEET.FE.ZS", 
                         "SL.UEM.NEET.MA.ZS")

```

## Define selected countries as vector

```{r}
#| eval: true

selected_countries_iso2c <- codelist |> 
  select(iso2c, iso.name.en) |> 
  filter(iso.name.en %in% c("Kenya", "Ghana", "Turkey", "Nepal")) |> 
  pull(iso2c)

```

## Download data

```{r}
#| cache: true

youth_unemployment <- WDI(indicator = selected_indicators, 
                          country = selected_countries_iso2c) |> 
  as_tibble()
```

## Tidy data

```{r}
youth_unemployment_tidy <- youth_unemployment  |> 
  rename(female = SL.UEM.NEET.FE.ZS,
         male = SL.UEM.NEET.MA.ZS) |> 
  pivot_longer(cols = female:male,
               names_to = "sex",
               values_to = "percent") |> 
  filter(!is.na(percent))
```

# Results & Discussion

## Youth Unemployment

```{r}
#| echo: false

country_name <- youth_unemployment_tidy |> 
  distinct(country) |> 
  pull()

min_max_year <- youth_unemployment_tidy |> 
  summarise(max = max(year),
            min = min(year))
```

Data was analysed for `r length(country_name)` countries: `r knitr::combine_words(country_name)`. The temporal data coverage is from `r pull(min_max_year, min)` to `r pull(min_max_year, max)`.

## Youth Unemployment

Unemployment rates are shown in @fig-unemployment.

```{r}
#| echo: false
#| fig-cap: "Youth unemployment rates in selected countries disaggregated by sex."
#| label: fig-unemployment

ggplot(data = youth_unemployment_tidy,
       aes(x = year, y = percent, color = sex)) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~country, nrow = 1) +
  theme_gray(base_size = 16) +
  scale_y_continuous(breaks = seq(0, 100, 20), limits = c(0, 90)) +
  theme(panel.grid.minor = element_blank(),
        legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 
```

## Youth Unemployment

Data for the latest available year is presented in @tbl-unemployment.

```{r}
#| echo: false
#| tbl-cap: "Youth unemployment rates (%) for the latest available year in selected countries and disaggregated by sex."
#| label: tbl-unemployment

youth_unemployment_tidy |> 
  group_by(country) |> 
  filter(year == max(year)) |> 
  select(-iso2c) |> 
  pivot_wider(names_from = sex,
              values_from = percent) |> 
  arrange(desc(female)) |> 
  knitr::kable(digits = 0) 
```

# References
