---
title: "Final Practice"
subtitle: "Section 03"
format: html
editor: visual
execute: 
  echo: true
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(here)
```

```{r}

treatment <- read_csv(here("wk-05/hw-05a-practice/data/what_a_waste/what-a-waste-city-level-data-treatment.csv"))

```

# Task 3.1

1.  Starting with the `treatment` data object, use functions taught during the course to answer the following question:

How many cities are covered in this dataset?

```{r}

treatment %>% 
  group_by(city) %>% 
  summarise(count = n()) %>% 
  summarise(count = n()) 


```

# Task 3.2

1.  Starting with the `treatment` data object, use functions taught during the course to answer the following question:

In how many cities is all waste (100%) openly dumped?

```{r}

treatment %>% 
  filter(waste_treatment == "open_dump" & percent == 100) %>%  
  summarise(count = n()) 


```

# Task 3.3

1.  Starting with the `treatment` data object, use functions taught during the course to answer the following question:

How many cities in which no waste (0%) is recycled fall into each income category?

```{r}

treatment %>% 
  filter(waste_treatment == "recycling" & percent == 0) %>% 
  group_by(income_id) %>% 
  summarise(count = n()) 


```

# Task 3.4

1.  Starting with the `treatment` data object, use functions taught during the course to answer the following question:

What is the mean, median, and standard deviation of treatment percentages for each waste treatment option and grouped by income?

```{r}

treatment %>% 
  group_by(waste_treatment, income_id) %>%
  summarise( 
    mean = mean(percent),
    median = median(percent), 
    sd = sd(percent)
  ) 


```

# Task 3.5

1.  In the YAML header, change the document format from `html` to `pdf`
2.  Render the document
3.  Open the 'Files' tab in the bottom right window
4.  Navigate to the generated PDF: `wk-05/hw-05a-practice/section-03.pdf`
5.  Tick the box next to the PDF file
6.  Click on 'More' (the blue gear icon) in the menu bar above the file manager
7.  Click on 'Export' and download the PDF file
8.  Send the PDF by email to the course instructor
